M.Sc. Proposal: \\ \bigskip Computational Electrokinetics

### 2013-02-01 notes/thesis-latex-template/thesis.tex
Update thesis plots.
Update thesis.

\setlength{\textheight}{9in}
\setlength{\textwidth}{7in}
\setlength{\topmargin}{0.05in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0.1in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\parindent}{0.0 in}
\setlength{\parskip}{0.1 in}

@@

The computational grid is created to have fine resolution near the particle surface, while
approximating the far-away boundary conditions. The logarithmic discretization is motivated
by the solution of the linearized system at Section \ref{app:linear}, 
where the largest gradients in the solution are near the particle surface.
The discretization of boundary conditions far away from the particle may be improved
by including the linear correction from the asymptotic solution, or applying

For strong electric field ($\beta \gg 1)$, the nonlinear effects become dominant and
another boundary-layer structure is expected to form near the particle surface. 
In order to achieve high numerical accuracy, the grid discretization should be made finer
near the particle surface (at $r = 1$), while $R_{max}$ can be made smaller.
In addition, the continuation of the solution between different $\beta$ values
should advance in finer steps, for Newton method convergence.

\begin{figure}
    \begin{center}
    \includegraphics[width=0.8\textwidth]{figs/Cubic/ionex_cubic1.eps}
    \includegraphics[width=0.8\textwidth]{figs/Cubic/ionex_cubic2.eps}
        \caption[Ion exchanger steady-state velocity]{
        Steady-state velocity $\cU$ as a function of the 
        electric field magnitude $\beta$, for $\gamma < 1$ (top) and $\gamma > 1$ (bottom). 
        The analytic cubic solution \eqref{eq:cubic} is represented by curves
        (solid for $\cU > 0$, dashed for $\cU < 0$), 
        and the numerical results are represented by markers 
        (circles for $\cU > 0$, squares for $\cU < 0$). Note that linear and
        cubic regimes match well the numerical results, including the critical $\beta$ 
        behaviour \eqref{eq:crit_beta}. 
        }
	    \label{fig:IonExCubic}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
	\includegraphics[width=0.8\textwidth]{figs/Cubic/Stream/N1.eps}
	\includegraphics[width=0.8\textwidth]{figs/Cubic/Stream/N2.eps}
	\includegraphics[width=0.8\textwidth]{figs/Cubic/Stream/N5.eps}
	\includegraphics[width=0.8\textwidth]{figs/Cubic/Stream/N10.eps}
        \caption[Ion exchanger streamlines -- numerical results]{
        Streamlines of the flow for various electric field $\beta$ values: 
        numerical results for $\gamma = 0.9$.  
        The steady-state velocity $\cU$ is positive, so the
        ion-exchanger drifts to the right. }
    \label{fig:Vortex1}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
	\includegraphics[width=0.8\textwidth]{figs/Cubic/Stream/A1.eps}
	\includegraphics[width=0.8\textwidth]{figs/Cubic/Stream/A2.eps}
	\includegraphics[width=0.8\textwidth]{figs/Cubic/Stream/A5.eps}
	\includegraphics[width=0.8\textwidth]{figs/Cubic/Stream/A10.eps}
        \caption[Ion exchanger streamlines -- analytical results]{
        Streamlines of the flow for various electric field $\beta$ values: 
        analytical approximation for $\gamma = 0.9$ (see Section \ref{sec:asymp}).
        The steady-state velocity $\cU$ is positive, so the
        ion-exchanger drifts to the right.}
    \label{fig:Vortex2}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/quadratic_order.eps}
        \caption[Quadratic convergence of steady-state velocity]{
        The discretization error between the numerical and analytical solutions for 
        $\alpha = 0$ ,$\Du = 0.5$, $\bar\zeta = 6$ and $R_{max} = 100$. The markers represent
        the discretization errors for various grid sizes, and the dotted lines correspond to linear
        functions of $\beta$, scaled by $4^n$. As mesh cell size decreases by 2, the discretization
        }
	    \label{fig:quadratic}
    \end{center}
\end{figure}

@@

>>>>

\setlength{\textheight}{9in}
\setlength{\textwidth}{7in}
\setlength{\topmargin}{0.05in}
\setlength{\headheight}{0.05in}
\setlength{\headsep}{0.1in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\parindent}{0.0 in}
\setlength{\parskip}{0.1 in}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/quadratic_order.eps}
        \caption[Quadratic convergence of steady-state velocity]{
        The discretization error between the numerical and analytical solutions for 
        $\alpha = 0$, $\Du = 0.5$, $\bar\zeta = 6$ and $R_{max} = 100$. The markers represent
        the discretization errors for various grid sizes, and the dotted lines correspond to linear
        functions of $\beta$, scaled by $4^n$. 
        The graph show that, as the number of grid points in each dimension increases by 2, 
        the discretization error decreases by 4.
        }
	    \label{fig:quadratic}
    \end{center}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Electrophoresis for large $\beta$ }
When the electric field is large, nonlinear effects dominate the solution.
In order to investigate the problem solution for strong electric fields, 
we have run the solver on the TAMNUN HPC cluster \cite{tamnun}.
Large grid sizes are used to verify that the numerical solution converges when the grid becomes finer and that no numerical artifacts are present in the solution.
Moreover, several $R_{max} \in \{10, 30, 100\}$ values are used, to make sure it has no major impact on the solution.
The results are shown at the following figures. 
The steady-state velocity results are shown for several grid resolutions and several $R_{max}$ values in Figure \ref{fig:LargeBetaV_grids}, and the estimated rate of convergence is shown at Figure \ref{fig:LargeBetaRatio}.

When $\beta$ becomes large, the salt concentration $C$ near the front of the particle 
(at $\theta \approx 0$) approaches zero, while creating a wake behind the particle 
(at $\theta \approx \pi$).
This is results in high gradients in $\log C$ near the front of the particle (Figures \ref{fig:LargeBeta_C1}, \ref{fig:LargeBeta_C2} and
\ref{fig:LargeBeta_C3}) and, therefore, strong electric field $\bE$ form near the particle surface (Figures \ref{fig:LargeBeta_Phi1}, \ref{fig:LargeBeta_Phi2} and \ref{fig:LargeBeta_Phi3}) due to effective boundary conditions at Eq. \ref{eq:ephor_bnd}. 
The streamline pattern doesn't change drastically for different electric field magnitudes, as shown in Figure \ref{fig:LargeBeta_Psi1}, \ref{fig:LargeBeta_Psi2} and \ref{fig:LargeBeta_Psi3}. 
The resulting flow pattern is similar to the Stokes creeping flow, as shown in Figure
\ref{fig:LargeBeta_PsiStokes}.

The numerical results suggest that a boundary 
layer starts to form around the front of the particle for strong
electric fields (see Figure \ref{fig:LargeBeta_BoundaryLayer}). 
This causes numerical problems to the current solver -- yielding a convergence
problems for large $\beta$. The convergence problems start earlier
for smaller grid size, and for higher $R_{max}$ values, probably
due to insufficient resolution at the $\brhat$ direction near
the particle surface.

\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_[128_256_512]_Rmax=10.eps}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_[512x512]_Rmax=10_30_100.eps}
        \caption[Large $\beta$ results for $\cU$]
{Large $\beta$ results for $\cU$ ($\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$). 
The left figure shows the numerical results of $\cU(\beta)$ for $R_{max} = 10$ and the following grid sizes: $128 \times 128$, $256 \times 256$, $512 \times 512$. 
The right figure shows the numerical results of $\cU(\beta)$ for
$512 \times 512$ grid for $R_{max} = 100, 30, 10$. For small values of $\beta$, the linear
regime of $\cU(\beta)$ is dominant. However, for larger values of $\beta$, the numerical solution
deviates significantly from the linear regime. For $\beta > 6$, the numerical solver convergence
is slowing and for $\beta \sim 7$, the solver fails to converge. As the grid resolution near
the particle surface becomes finer (when $N_r$ increases or $R_{max}$ decreases), the solver
convergence is better.}
	    \label{fig:LargeBetaV_grids}	    
    \end{center}
\end{figure}
\begin{figure}
    \begin{center}
    \includegraphics[width=.5\textwidth]{figs/LargeBetaRatio.eps}
        \caption[Rate of convergence estimation for $\cU$ at large $\beta$]
        {Rate of convergence estimation for $\cU$ at large $\beta$  
        ($\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$). 
        The numerical results of $\cU(\beta)$, for the grid sizes of 
        $128 \times 128$, $256 \times 256$, $512 \times 512$. 
        The rate of convergence $r(\beta)$ is estimated by 
        $r(\beta) \triangleq \left(\cU^{4h}(\beta) - \cU^{2h}(\beta)\right)/\left(\cU^{2h}(\beta) - \cU^{h}(\beta)\right)$.
        For small $\beta$, $r(\beta) \rightarrow 4$, showing 
        quadratic convergence for the linear regime. 
        For moderate $\beta$, the convergence rate slows down 
        and deteriorates for large $\beta$ values.
        }
	    \label{fig:LargeBetaRatio}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LargeBeta/sol_beta=1.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_C.eps}
        \caption[$\beta = 0.1$ results for $C$]
        {$\beta = 0.1$ results for $C$: grid size $512 \times 512$, 
        $R_{max} = 10$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        Here the linear regime dominates the solution for 
        $C_\beta(r, \theta) \approx 1 - \beta \frac{\cos\theta}{2r^2}$.}
	    \label{fig:LargeBeta_C1}	    
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LargeBeta/sol_beta=8.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_C.eps}
        \caption[$\beta = 0.8$ results for $C$]
        {$\beta = 0.8$ results for $C$: grid size $512 \times 512$, 
        $R_{max} = 10$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        For moderate $\beta$ values, the salt concentration starts
        to form a wake behind the particle ($\theta \approx \pi$).}
	    \label{fig:LargeBeta_C2}	    
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LargeBeta/sol_beta=5.000e+00_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_C.eps}
        \caption[$\beta = 5.0$ results for $C$]
        {$\beta = 5.0$ results for $C$: grid size $512 \times 512$, 
        $R_{max} = 10$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        For large values of $\beta$, there is a significant drop in
        salt concentration at the front of the particle -- 
        $C$ approaches zero at $\theta \approx 0$, such that 
        $\deriv{C}{r}$ becomes large 
        (since the change in $C$ happens at a small length scale).
        Due to the boundary conditions of the electrophoresis problem
        (given by Eq. \ref{eq:ephor_bnd}), the electric field radial
        component given by 
        $E_r = -\deriv{\varPhi}{r} = -\frac{1}{C}\deriv{C}{r}$, 
        becomes singular where $C \rightarrow 0$.
        }
	    \label{fig:LargeBeta_C3}	    
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LargeBeta/sol_beta=1.000e-01_[129x129]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_PhiE.eps}
        \caption[$\beta = 0.1$ results for $\varPhi$]
        {$\beta = 0.1$ results for $\varPhi$:
        grid size $512 \times 512$, $R_{max} = 10$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$. 
        The electric potential and the electric field results are multiplied by $\beta^{-1}$,
	    so the linear regime appears unchanged for different values of $\beta$.
        Here the linear regime dominates the solution for 
        $\varPhi_\beta(r, \theta) \approx -\beta {r \cos\theta}$,
        hence the electric field is constant for weak electric fields:
        $\bE \approx \beta \ui$.}
	    \label{fig:LargeBeta_Phi1}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LargeBeta/sol_beta=8.000e-01_[129x129]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_PhiE.eps}
        \caption[$\beta = 0.8$ results for $\varPhi$]
        {$\beta = 0.8$ results for $\varPhi$:
        grid size $512 \times 512$, $R_{max} = 10$, $\Du = 1$, 
        $\bar\zeta = 10$, $\alpha = 0.5$. 
	    The electric potential and the electric field results are multiplied by $\beta^{-1}$,
	    so the linear regime appears unchanged for different values of $\beta$.
        For moderate values of $\beta$,
        the symmetry breaks and the electric field becomes stronger 
        near the front of the particle than at the back.}
	    \label{fig:LargeBeta_Phi2}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LargeBeta/sol_beta=5.000e+00_[129x129]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_PhiE.eps}
        \caption[$\beta = 5.0$ results for $\varPhi$]
        {$\beta = 5.0$ results for $\varPhi$:
        grid size $512 \times 512$, $R_{max} = 10$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The electric potential and the electric field results are multiplied by $\beta^{-1}$,
	    so the linear regime appears unchanged for different values of $\beta$.
	    For large values of $\beta$, the electric field radial component $E_r = -\deriv{\varPhi}{r}$
	    at the particle front becomes much larger than the electric field at the particle back.
		}
	    \label{fig:LargeBeta_Phi3}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LargeBeta/sol_beta=1.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
        \caption[$\beta = 0.1$ results for $\Psi$]
        {$\beta = 0.1$ streamlines results: grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The numerical results for $\Psi$ is scaled by $\beta^{-1}$, so the linear regime
        will appear the same for different $\beta$ values.}
	    \label{fig:LargeBeta_Psi1}	    
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LargeBeta/sol_beta=8.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
        \caption[$\beta = 0.8$ results for $\Psi$]
        {$\beta = 0.8$ streamlines results: grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The numerical results for $\Psi$ is scaled by $\beta^{-1}$, so the linear regime
        will appear the same for different $\beta$ values.}
	    \label{fig:LargeBeta_Psi2}	    
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LargeBeta/sol_beta=5.000e+00_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
        \caption[$\beta = 5.0$ results for $\Psi$]
        {$\beta = 5.0$ streamlines results: grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The numerical results for $\Psi$ is scaled by $\beta^{-1}$, so the linear regime
        will appear the same for different $\beta$ values.}
	    \label{fig:LargeBeta_Psi3}	    
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/LinearStokes_Psi.eps}
        \caption[Stokes flow results]
        {Stokes flow streamlines results: grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The numerical results for $\Psi$ is scaled by $\beta^{-1}$, so the linear regime
        will appear the same for different $\beta$ values. Note that the numerical results
        show very similar flow pattern to the analytical Stokes flow.}
	    \label{fig:LargeBeta_PsiStokes}	    
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/BoundaryLayerWidth.eps}
        \caption[Electric field boundary layer]
        {Electric field boundary layer formation: grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$. The left
        graph shows normalized electric field radial component $E_r$ near the particle front
        ($\theta \approx 0$). The numerical results for $e(\rho) = E_r(r=1+\rho, \theta=0)$
        are normalized by $\hat{e}(\rho) = \left(e(\rho) - \beta\right)/\left(e(0) - \beta\right)$,
        where $e(\rho = \infty) = \beta$.
        The normalized electric field satisfies $\hat{e}(\rho) \in [0, 1]$ and is shown as function
        of $\rho = r - 1$ at the left graph.
        In order to estimate the width of the boundary layer, for each $\beta$, we compute the value 
        of $\bar\rho(\beta) = \min_\rho \{ \rho : \hat{e}(\rho) < 0.5\}$. Each such $\bar\rho(\beta)$
        is plotted at the left graph as a solid dot on the corresponding line.
        In addition, $\bar\rho(\beta)$ is plotted as a function of $\beta$ using logarithmic 
        $y$ axis at the right graph to show that $\bar\rho(\beta) \propto e^{-A\beta}$.
        Thus, a boundary layer is indeed forming near the front of the particle.
        }
	    \label{fig:LargeBeta_BoundaryLayer}	    
    \end{center}
\end{figure}

The computational grid is created to have fine resolution near the particle surface, while
approximating the far-away boundary conditions. The logarithmic discretization is motivated
by the solution of the linearized system at Section \ref{app:linear}, 
where the largest gradients in the solution are near the particle surface.
The discretization of boundary conditions far away from the particle may be improved
by including the linear correction from the asymptotic solution, or reformulation of
far-away boundary condition according to the asymptotic behavior of the analytic solution
far away from the particle.

For strong electric field ($\beta \gg 1)$, the nonlinear effects become dominant and
another boundary-layer structure begins to form near the particle surface. 
In order to achieve high numerical accuracy, the grid discretization should be made finer
near the particle surface (at $r = 1$), while $R_{max}$ can be made smaller. Moreover, different discretization scheme may be used for
the $r$ coordinate, to have higher grid resolution near the particle.
In addition, the continuation of the solution between different $\beta$ values
should advance in finer steps, to ensure Newton method convergence.
